---
title: "Community Data Science meets Digital Heritage: Controversies about World Heritage Sites in the Online Community of Wikipedia"
author:
  - Ben Marwick
  - Prema Smith
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
---


<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)
```



# Introduction

Heritage is the processes and outcomes of people engaging with elements of the past – material and immaterial – and attributing social and cultural meanings to them in the present (Smith 2006; Harrison 2013). These are important to understand because they shape peoples' identidies and influence how they think and behave towards other people. Digital heritage are these engagements that are enabled by the Internet (Bonacchi and Krzyzanska 2019), leaving traces that can be identified and quantified using data science methods. Recent work on digital heritage has opened a new field of study.s


• Represents a major turn from post- modernism, critical theory, & deeply qualitative, towards digital heritage ontologies, data-intensive ethnographies, role of heritage scholars

# Background

# Methods

# Results

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at?
git2r::repository(here::here())
```
